# Visualization with drake {#vis}

```{r, echo = FALSE}
suppressMessages(suppressWarnings(library(drake)))
knitr::opts_chunk$set(
  collapse = TRUE,
  error = TRUE,
  warning = TRUE
)
```

`Drake` has powerful visuals to help you plan your project. You can generate an interactive workflow network with either `drake_graph()` or `vis_drake_graph()`. Then click, drag, hover, zoom, and pan. Use either the mouse or the green buttons near the bottom.

## Dependency reactivity

Initially, your entire project is out of date.

```{r graphoutdated}
library(drake)
load_mtcars_example() # Get the code with drake_example("mtcars").
config <- drake_config(my_plan)
vis_drake_graph(config) # Same as drake_graph()
```

After `make()`, the whole project is all caught up.

```{r graphmake}
config <- make(my_plan, jobs = 4, verbose = FALSE)
vis_drake_graph(config)
```


But when you change a dependency, some targets are out of date until the next `make(my_plan)`.

```{r reg2graphvisual}
reg2 <- function(d){
  d$x3 <- d$x ^ 3
  lm(y ~ x3, data = d)
}
vis_drake_graph(config)
```

## Subgraphs

Graphs can grow enormous for serious projects, so there are multiple ways to focus on a manageable subgraph. The most brute-force way is to just pick a manual `subset` of nodes. However, with the `subset` argument, `vis_drake_graph()` may drop intermediate nodes and edges.

```{r subsetgraph}
vis_drake_graph(
  config,
  subset = c("regression2_small", file_store("report.md"))
)
```

The rest of the subgraph functionality preserves connectedness. Use `targets_only` to ignore the imports.

```{r targetsonly}
vis_drake_graph(config, targets_only = TRUE)
```

Similarly, you can just show downstream nodes.

```{r fromout}
vis_drake_graph(config, from = c("regression2_small", "regression2_large"))
```

Or upstream ones.

```{r fromin}
vis_drake_graph(config, from = "small", mode = "in")
```

In fact, let us just take a small neighborhood around a target in both directions.


```{r fromall}
vis_drake_graph(config, from = "small", mode = "all", order = 1)
```

## Control the legend.

To remove superfluous information from the legend, set the `full_legend` argument to `FALSE`.

```{r smalllegend}
vis_drake_graph(config, full_legend = FALSE)
```

To remove the legend altogether, set the `ncol_legend` argument to `0`.

```{r nolegend}
vis_drake_graph(config, ncol_legend = 0)
```

## More flexibility

We have only scratched the surface of `vis_drake_graph()`. The help files (`?vis_drake_graph`) document much more functionality. In particular, the `dataframes_graph()` and `render_drake_graph()` functions let you customize your own [visNetwork graph](http://datastorm-open.github.io/visNetwork/).
