# Dynamic branching {#dynamic}

```{r, message = FALSE, warning = FALSE, echo = FALSE}
knitr::opts_knit$set(root.dir = fs::dir_create(tempfile()))
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(broom)
library(drake)
library(gapminder)
library(tidyverse)
```

## A note about versions

The first release of dynamic branching was in `drake` version 7.8.0. In subsequent versions, dynamic branching behaves differently. This manual describes how dynamic branching works in development `drake` (to become version 7.9.0 in early January 2020). If you are using version 7.8.0, please refer to [this version of the chapter](https://github.com/ropensci-books/drake/blob/c4dfa6dd71b5ffa4c6027633ae048d2ab0513c6d/dynamic.Rmd) instead.

## Motivation

In large workflows, you may need more targets than you can easily type in a plan, and you may not be able to fully specify all targets in advance. Dynamic branching is an interface to declare new targets while `make()` is running. It lets you create more compact plans and graphs, it is easier to use than [static branching](#static), and it improves the startup speed of `make()` and friends.

## Which kind of branching should I use?

With dynamic branching, `make()` is faster to initialize, and you have far more flexibility. With [static branching](#static), you have meaningful target names, and it is easier to predict what the plan is going to do in advance. There is a ton of room for overlap and personal judgement, and you can even use both kinds of branching together.

## Dynamic targets

A dynamic target is a [vector](https://vctrs.r-lib.org/) of *sub-targets*. We let `make()` figure out which sub-targets to create and how to aggregate them.

As an example, let's fit a regression model to each continent in [Gapminder data](https://github.com/jennybc/gapminder). To activate dynamic branching, use the `dynamic` argument of `target()`.

```{r}
library(broom)
library(drake)
library(gapminder)
library(tidyverse)

# Split the Gapminder data by continent.
gapminder_continents <- function() {
  gapminder %>%
    mutate(gdpPercap = scale(gdpPercap)) %>%
    split(f = .$continent)
}

# Fit a model to a continent.
fit_model <- function(continent_data) {
  data <- continent_data[[1]]
  data %>%
    lm(formula = gdpPercap ~ year) %>%
    tidy() %>%
    mutate(continent = data$continent[1]) %>%
    select(continent, term, statistic, p.value)
}

plan <- drake_plan(
  continents = gapminder_continents(),
  model = target(fit_model(continents), dynamic = map(continents))
)

make(plan)
```

Thanks to [`vctrs`](https://vctrs.r-lib.org/), the data type of every sub-target is the same as the dynamic target it belongs to. In other words, `model` and `model_23022788` are both data frames. To understand how `drake` splits and concatenates dynamic targets, see functions [`vec_size()`](https://vctrs.r-lib.org/reference/vec_size.html),   [`vec_slice()`](https://vctrs.r-lib.org/reference/vec_slice.html), and [`vec_c()`](https://vctrs.r-lib.org/reference/vec_c.html).

```{r}
readd(model)
```

If automatic concatenation of sub-targets is confusing (e.g. if some sub-targets are `NULL`, as in <https://github.com/ropensci-books/drake/issues/142>) you can read the dynamic target as a named list (only in `drake` version 7.10.0 and above).

```{r}
readd(model, subtarget_list = TRUE) # Requires drake >= 7.10.0.
```

Alternatively, you can work with the individual sub-targets.

```{r}
subtargets(model)

readd(model_23022788) # equivalent: readd(model, subtargets = 1)
```

But automatic concatenation with `vctrs` is the default behavior, and it is how `drake` thinks about downstream targets. In our example, when we write `dynamic = map(continents)` in the plan, `drake` treats `continents` as a [vector](https://vctrs.r-lib.org/). In other words, `target(fit_model(continents), dynamic = map(continents))` is equivalent to commands `fit_model(continents[1])` through `fit_model(continents[5])`. Since `continents` is really a list of data frames, `continents[1]` through `continents[5]` are also lists of data frames, which is why we need the line `data <- continent_data[[1]]` in `fit_model()`. This may seem inconvenient at first, but it comes with extra type stability, and it allows us to reference a dynamic target and its sub-targets interchangeably.

To post-process our models, we can work with either the individual sub-targets or the whole vector of all the models. Below, `year` uses the former and `intercept` uses the latter.

```{r}
plan <- drake_plan(
  continents = gapminder_continents(),
  model = target(fit_model(continents), dynamic = map(continents)),
  # Filter each model individually:
  year = target(filter(model, term == "year"), dynamic = map(model)),
  # Aggregate all the models, then filter the whole vector:
  intercept = filter(model, term != "year")
)

make(plan)
```

```{r}
readd(year)
```

```{r}
readd(intercept)
```

## Dynamic transformations

Dynamic branching supports transformations `map()`, `cross()`, and `group()`. These transformations tell `drake` how to create sub-targets.

### `map()`

`map()` iterates over the [vector slices](https://vctrs.r-lib.org/reference/vec_slice.html) of the targets you supply as arguments. We saw above how `map()` iterates over lists. If you give it a data frame, it will map over the rows.

```{r}
plan <- drake_plan(
  subset = head(gapminder),
  row = target(subset, dynamic = map(subset))
)

make(plan)
```

```{r}
readd(row_9939cae3)
```

If you supply multiple targets, `map()` iterates over the slices of each.

```{r}
plan <- drake_plan(
  numbers = seq_len(2),
  letters = c("a", "b"),
  zipped = target(paste0(numbers, letters), dynamic = map(numbers, letters))
)

make(plan)
```

```{r}
readd(zipped)
```

### `cross()`

`cross()` creates a new sub-target for each combination of targets you supply as arguments.

```{r}
plan <- drake_plan(
  numbers = seq_len(2),
  letters = c("a", "b"),
  combo = target(paste0(numbers, letters), dynamic = cross(numbers, letters))
)

make(plan)
```

```{r}
readd(combo)
```

### `group()`

With `group()`, you can create multiple aggregates of a given target. Use the `.by` argument to set a grouping variable.

```{r}
plan <- drake_plan(
  data = gapminder,
  by = data$continent,
  gdp = target(
    tibble(median = median(data$gdpPercap), continent = by[1]),
    dynamic = group(data, .by = by)
  )
)

make(plan)
```

```{r}
readd(gdp)
```

## Trace

All dynamic transforms have a `.trace` argument to record optional metadata for each sub-target. In the example from `group()`, the trace is another way to keep track of the continent of each median GDP value.

```{r}
plan <- drake_plan(
  data = gapminder,
  by = data$continent,
  gdp = target(
    median(data$gdpPercap),
    dynamic = group(data, .by = by, .trace = by)
  )
)

make(plan)
```

The `gdp` target no longer contains any explicit reference to continent.

```{r}
readd(gdp)
```

However, we can look up the continents in the trace.

```{r}
read_trace("by", gdp)
```


## `max_expand`

Suppose we want a model for each *country*.

```{r}
gapminder_countries <- function() {
  gapminder %>%
    mutate(gdpPercap = scale(gdpPercap)) %>%
    split(f = .$country)
}

plan <- drake_plan(
  countries = gapminder_countries(),
  model = target(fit_model(countries), dynamic = map(countries))
)
```

The Gapminder dataset has 142 countries, which can get overwhelming. In the early stages of the workflow when we are still debugging and testing, we can limit the number of sub-targets using the `max_expand` argument of `make()`.

```{r}
make(plan, max_expand = 2)
```

```{r}
readd(model)
```

Then, when we are confident and ready, we can scale up to the full number of models.

```{r, eval = FALSE}
make(plan)
```
