# Dynamic branching {#dynamic}

```{r, message = FALSE, warning = FALSE, echo = FALSE}
knitr::opts_knit$set(root.dir = fs::dir_create(tempfile()))
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(broom)
library(drake)
library(gapminder)
library(tidyverse)
```

## Motivation

With [static branching](#static) (explained in the [following chapter](#static)) we can concisely create plans with large numbers of targets. However, static branching has major issues.

1. If a plan gets too large, functions `drake_config()` and `outdated()` become very slow, which severely slows down the initialization of `make()`.
2. We need to declare every single target in advance. We cannot *define* new targets based the values returned from other targets.
3. The [graph visualizations](#visuals) are too slow, too cumbersome, and too unresponsive when the number of static targets is large.
4. [static branching](#static) relies on metaprogramming and code manipulation, which makes it difficult to use and understand.

[Dynamic branching](#dynamic), supported in versions above 7.7.0, solves these problems.

## Which kind of branching should I use?

With dynamic branching, `make()` is faster to initialize, and you have far more flexibility. With [static branching](#static), you have meaningful target names, and it is easier to predict what the plan is going to do in advance. There is a ton of room for overlap and personal judgement, and you can even use both kinds of branching together!

```{r}
library(drake)

plan <- drake_plan(
  locations = get_location_db(), # many rows
  coord = target(
    geocode_location(locations, method),
    dynamic = map(locations),
    transform = map(method = c("address", "postal", "boundary")) #static
  ),
  results = target(
    bind_rows(coord),
    dynamic = group(coord),
    transform = combine(coord)
  )
)

plot(plan)
```

## Dynamic targets

A dynamic target has multiple *sub-targets*. Prior to running `make()`, we do not know how many sub-targets there will be, nor what they will contain. This flexibility lets the data drive the plan. For example, we can fit a regression model to each continent in [Gapminder data](https://github.com/jennybc/gapminder) and give each model its own target. To activate dynamic branching, use the `dynamic` argument of `target()`.

```{r}
library(broom)
library(drake)
library(gapminder)
library(tidyverse)

fit_model <- function(dataset, continent) {
  dataset %>%
    filter(continent == !!continent) %>% # The !! is important.
    lm(formula = gdpPercap ~ year) %>%
    tidy() %>%
    mutate(continent = !!continent)
}

plan <- drake_plan(
  # This dataset can change, and we want
  # the downstream targets to update.
  dataset = gapminder %>%
    mutate(gdpPercap = scale(gdpPercap)),
  
  # We need a grouping variable.
  continent = unique(dataset$continent),
  
  # Fit GDP vs year for each country.
  model = target(
    fit_model(dataset, continent),
    dynamic = map(continent) # Activate dynamic branching!
  )
)

make(plan)
```

The sub-targets have strange names ([there are good reasons!](https://github.com/ropensci/drake/issues/685#issuecomment-549096373)) but you do not need sub-target names in order to fetch values.

```{r}
readd(model, subtargets = c(1, 2))
```

To select specific targets, simply load the original grouping variable and select the indices you need.

```{r}
loadd(continent)
index <- which(continent == "Oceania")
readd(model, subtargets = index)[[1]]
```

The visuals load faster and look nicer because we omit the sub-targets.

```{r}
config <- drake_config(plan)
vis_drake_graph(config)
```

## map()

The dynamic `map()` transformation creates a new sub-target for each element of the grouping variables you supply. 

```{r}
plan <- drake_plan(
  static_numbers = seq_len(2),
  static_letters = c("a", "b"),
  dynamic_lowercase = target(
    paste0(static_numbers, static_letters),
    dynamic = map(static_numbers, static_letters)
  ),
  dynamic_uppercase = target(
    toupper(dynamic_lowercase),
    dynamic = map(dynamic_lowercase)
  )
)

make(plan)
```

For grouping variables that are vectors and simple lists, dynamic branching iterates over the elements.

```{r}
readd(dynamic_lowercase)

readd(dynamic_uppercase)
```

For grouping variables that are array-like objects (anything with a non-null `dim()`) dynamic branching iterates over the *rows*. This includes not only arrays and matrices, but *data frames* and *tibble*s as well.[^1]

[^1]: This behavior is a deliberate design choice. Yes, it contradicts `purrr::map()`, but [row-oriented workflows](https://github.com/jennybc/row-oriented-workflows) come up far more often than column-oriented workflows in `drake`. If you want to loop over the columns of a data frame, convert it to a list first.

```{r}
plan <- drake_plan(
  dataset = head(gapminder, n = 3),
  row = target(dataset, dynamic = map(dataset))
)

make(plan)

readd(row)
```

Regardless of the data type of object `x`, `NROW(x)` tells you how many sub-targets you will get from `map(x)`.

```{r}
NROW(letters)

NROW(mtcars)
```


## cross()

`cross()` is like `map()` except we create a new target for each combination of grouping variables.


```{r}
plan <- drake_plan(
  numbers = seq_len(2),
  letters = c("a", "b"),
  result = target(
    c(numbers, letters),
    dynamic = cross(numbers, letters)
  )
)

make(plan)

readd(result)
```

## `group()`

`group()` covers both splitting and aggregation. It chops up static targets into multiple pieces, and it gathers together existing sub-targets from dynamic dependencies. In both cases, the `.by` argument controls the aggregation. Let's fit a model to each continent in the Gapminder dataset and then combine all the results at the end.

```{r}
fit_model <- function(dataset) {
  dataset %>%
    lm(formula = gdpPercap ~ year) %>%
    tidy() %>%
    mutate(continent = dataset$continent[1])
}
  
plan <- drake_plan(
  # Let's fit a model for each continent and then
  # combine the results at the end.
  dataset = gapminder %>%
    mutate(gdpPercap = scale(gdpPercap)),
  
  # We need a target to act as a grouping variable.
  continent = dataset$continent,
  
  # Static data splitting: fit a model for each continent.
  model = target(
    fit_model(dataset),
    dynamic = group(dataset, .by = continent)
  ),
  
  # Dynamic data aggregation: bring the models together.
  results = target(
    bind_rows(model),
    # Ignore .by to get a single all-encompassing group.
    dynamic = group(model)
  )
)

make(plan)

readd(results)[[1]]
```

## Tracing sub-dependencies

In dynamic branching, tracing keeps track of the sub-dependencies that correspond to individual the sub-targets. We can use tracing to remember which continent belongs to which model. `map()`, `cross()`, and `group()` each support a `.trace` argument for this purpose.

```{r}
fit_model <- function(dataset, continent) {
  dataset %>%
    filter(continent == !!continent) %>%
    lm(formula = gdpPercap ~ year) %>%
    tidy() %>%
    # The continent we put here should match the trace.
    mutate(continent = !!continent)
}

plan <- drake_plan(
  dataset = gapminder %>%
    mutate(gdpPercap = scale(gdpPercap)),
  continent = as.character(unique(dataset$continent)),
  model = target(
    fit_model(dataset, continent),
    dynamic = map(
      continent,
      # Use the trace to remember the continent of each model.
      .trace = continent
    )
  )
)

make(plan)
```

Now, we can use `read_trace()` to find out which continent belongs to each model. As you can see, the trace matches up with the continents we identify in the model output itself.

```{r}
read_trace("continent", "model")[seq_len(3)]

readd(model, subtargets = seq_len(3))
```

The trace also helps us create grouping variables for dynamic targets downstream of `cross()` and `group()`. Below, we fit multiple models to multiple countries and aggregate on country. `get_trace()` gets grouping variables easily.

```{r}
# We need a function to fit the model.
fit_model <- function(dataset, country, covariate) {
  formula <- paste0("gdpPercap ~ ", covariate) %>%
    as.formula()
  dataset %>%
    filter(country == !!country) %>% # The !! is important.
    lm(formula = formula) %>%
    tidy() %>%
    mutate(covariate = covariate, country = !!country)
}

plan <- drake_plan(
  # This dataset can change, and we want
  # the downstream targets to update.
  dataset = gapminder %>%
    mutate(gdpPercap = scale(gdpPercap)) %>%
    mutate(pop = scale(pop)),
  
  # We need a grouping variable for country,
  country = unique(as.character(dataset$country)),
  
  # and another for the covariate
  covariate = c("pop", "year"),
  
  # Fit GDP for each country and covariate.
  model = target(
    fit_model(dataset, country, covariate),
    dynamic = cross(
      country,
      covariate,
      # Use the trace so we can get each model's
      # country and covariate later on.
      .trace = c(country, covariate)
    )
  ),
  
  # Get the country of each model.
  country_trace = get_trace("country", model),
  
  # Aggregate the models by country
  models_by_country = target(
    bind_rows(model),
    dynamic = group(
      model,
      .by = country_trace,
      # Define a new trace on the aggregates
      # so we can look up the countries later.
      .trace = country_trace
    )
  )
)

make(plan, verbose = 0L)
```

As before, notice how the trace matches the countries in the model output.

```{r}
read_trace("country_trace", "models_by_country")[seq_len(3)]

readd(models_by_country, subtargets = seq_len(3))
```

## max_expand

If we want to fit a model for every single *country* in the Gapminder dataset, the number of sub-targets grows large, and the workflow becomes cumbersome.

```{r}
length(unique(gapminder$country)) # number of countries
```

To make things easier to manage, we should develop and test on a downsized version of the workflow before scaling up. Simply set `max_expand` to a small number in `make()`. That way, `drake` only builds the first few sub-targets for each dynamic target.

```{r}
library(broom)
library(drake)
library(gapminder)
library(tidyverse)

fit_model <- function(dataset, country) {
  dataset %>%
    filter(country == !!country) %>% # The !! is important.
    lm(formula = gdpPercap ~ year) %>%
    tidy() %>%
    mutate(country = !!country)
}

plan <- drake_plan(
  # This dataset can change, and we want
  # the downstream targets to update.
  dataset = gapminder %>%
    mutate(gdpPercap = scale(gdpPercap)),
  
  # We need a grouping variable.
  country = unique(dataset$country),
  
  # Fit GDP vs year for each country.
  model = target(
    fit_model(dataset, country),
    dynamic = map(country) # Activate dynamic branching!
  )
)

make(plan, max_expand = 2) # Only run the first 2 models.

readd(model)
```

If we scale up to more models and leave the dependencies alone, the old models stay up to date.

```{r}
make(plan, max_expand = 3) # previously 2

readd(model)
```

Likewise, we can reuse old models if we scale back down.

```{r}
make(plan, max_expand = 1) # Does not fit any new models.

# All make() did was associate the dynamic target "model" with
# a different set of sub-targets.
progress()

readd(model)
```
