# Dynamic branching {#dynamic}

```{r, message = FALSE, warning = FALSE, echo = FALSE}
knitr::opts_knit$set(root.dir = fs::dir_create(tempfile()))
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(broom)
library(drake)
library(gapminder)
library(tidyverse)
```

## Motivation

With [static branching](#static) (explained in the [following chapter](#static)) we can concisely create plans with large numbers of targets. However, static branching has major issues.

1. If a plan gets too large, functions `drake_config()` and `outdated()` become very slow, which severely slows down the initialization of `make()`.
2. We need to declare every single target in advance. We cannot *define* new targets based the values returned from other targets.
3. The [graph visualizations](#visuals) are too slow, too cumbersome, and too unresponsive when the number of static targets is large.
4. [static branching](#static) relies on metaprogramming and code manipulation, which makes it difficult to use and understand.

[Dynamic branching](#dynamic), supported in versions above 7.7.0, solves these problems.

## Which kind of branching should I use?

With dynamic branching, `make()` is faster to initialize, and you have far more flexibility. With [static branching](#static), you have meaningful target names, and it is easier to predict what the plan is going to do in advance. There is a ton of room for overlap and personal judgement, and you can even use both kinds of branching together!

```{r}
library(drake)

plan <- drake_plan(
  locations = get_location_db(), # many rows
  coordinates = target(
    my_geocode_fn(locations, method),
    dynamic = map(locations),
    transform = map(method = c("address", "postal", "boundary")) #static
  ),
  results = target(
    bind_rows(coordinates),
    dynamic = combine(coordinates),
    transform = combine(coordinates)
  )
)

plot(plan)
```

## Dynamic targets

A dynamic target has multiple *sub-targets*. Prior to running `make()`, we do not know how many sub-targets there will be, nor what they will contain. This flexibility lets the data drive the plan. For example, we can fit a regression model to each continent in [Gapminder data](https://github.com/jennybc/gapminder) and give each model its own target. To activate dynamic branching, use the `dynamic` argument of `target()`.

```{r}
library(broom)
library(drake)
library(gapminder)
library(tidyverse)

fit_model <- function(dataset, continent) {
  dataset %>%
    filter(continent == !!continent) %>% # The !! is important.
    lm(formula = gdpPercap ~ year) %>%
    tidy() %>%
    mutate(continent = !!continent)
}

plan <- drake_plan(
  # This dataset can change, and we want
  # the downstream targets to update.
  dataset = gapminder %>%
    mutate(gdpPercap = scale(gdpPercap)),
  
  # We need a grouping variable .
  continent = unique(dataset$continent),
  
  # Fit GDP vs year for each country.
  model = target(
    fit_model(dataset, continent),
    dynamic = map(continent) # Activate dynamic branching!
  )
)

make(plan)
```

The sub-targets have strange names ([there are good reasons!](https://github.com/ropensci/drake/issues/685#issuecomment-549096373)) but you do not need sub-target names in order to fetch values.

```{r}
readd(model, subtargets = c(1, 2))
```

To select specific targets, simply load the original grouping variable and select the indices you need.

```{r}
loadd(continent)
index <- which(continent == "Oceania")
readd(model, subtargets = index)[[1]]
```

The visuals load faster and look nicer because we omit the sub-targets.

```{r}
config <- drake_config(plan)
vis_drake_graph(config)
```

## map()

The dynamic `map()` transformation creates a new sub-target for each element of the grouping variables you supply. 

```{r}
plan <- drake_plan(
  static_numbers = seq_len(2),
  static_letters = c("a", "b"),
  dynamic_lowercase = target(
    paste0(static_numbers, static_letters),
    dynamic = map(static_numbers, static_letters)
  ),
  dynamic_uppercase = target(
    toupper(dynamic_lowercase),
    dynamic = map(dynamic_lowercase)
  )
)

make(plan)
```

For grouping variables that are vectors and simple lists, dynamic branching iterates over the elements.

```{r}
readd(dynamic_lowercase)

readd(dynamic_uppercase)
```

For grouping variables that are array-like objects (anything with a non-null `dim()`) dynamic branching iterates over the *rows*. This includes not only arrays and matrices, but *data frames* and *tibble*s as well.[^1]

[^1]: This behavior is a deliberate design choice. Yes, it contradicts `purrr::map()`, but [row-oriented workflows](https://github.com/jennybc/row-oriented-workflows) come up far more often than column-oriented workflows in `drake`. If you want to loop over the columns of a data frame, convert it to a list first.

```{r}
plan <- drake_plan(
  dataset = head(gapminder, n = 3),
  row = target(dataset, dynamic = map(dataset))
)

make(plan)

readd(row)
```

Regardless of the data type of object `x`, `NROW(x)` tells you how many sub-targets you will get from `map(x)`.

```{r}
NROW(letters)

NROW(mtcars)
```


## cross()

`cross()` is like `map()` except we create a new target for each combination of grouping variables.


```{r}
plan <- drake_plan(
  numbers = seq_len(2),
  letters = c("a", "b"),
  result = target(
    c(numbers, letters),
    dynamic = cross(numbers, letters)
  )
)

make(plan)

readd(result)
```

## `combine()`

`combine()` can group together sub-targets or split up static targets. The `.by` argument lets us control the aggregation. Let's fit a model to each continent in the Gapminder dataset and then combine all the results at the end.

```{r}
fit_model <- function(dataset) {
  dataset %>%
    lm(formula = gdpPercap ~ year) %>%
    tidy() %>%
    mutate(continent = dataset$continent[1])
}
  
plan <- drake_plan(
  # Let's fit a model for each continent and then
  # combine the results at the end.
  dataset = gapminder %>%
    mutate(gdpPercap = scale(gdpPercap)),
  
  # We need a target to act as a grouping variable.
  continent = dataset$continent,
  
  # Fit a model for each continent.
  model = target(
    fit_model(dataset),
    dynamic = combine(dataset, .by = continent)
  ),
  
  # Aggregate the results together.
  results = target(
    bind_rows(model),
    dynamic = combine(model) # no .by necessary
  )
)

make(plan)

readd(results)[[1]]
```