# Finding the best model of gross state product {#gsp}

```{r, echo = F}
suppressPackageStartupMessages(library(biglm))
suppressPackageStartupMessages(library(drake))
suppressPackageStartupMessages(library(Ecdat))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(purrr))
suppressPackageStartupMessages(library(tidyverse))
unlink(".drake", recursive = TRUE)
clean(destroy = TRUE, verbose = FALSE)
unlink(
  c("gsp", "Makefile", "report.Rmd", "shell.sh", "STDIN.o*", "Thumbs.db"),
  recursive = TRUE
)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

The following data analysis workflow shows off `drake`'s ability to generate lots of reproducibly-tracked tasks with ease. The same technique would be cumbersome, even intractable, with [GNU Make](https://www.gnu.org/software/make/).

## Get the code.

Write the code files to your workspace.

```{r getthequickstartcode03, eval = FALSE}
drake_example("gsp")
```

The new `gsp` folder now includes a file structure of a serious `drake` project, plus an `interactive-tutorial.R` to narrate the example. The code is also [online here](https://github.com/wlandau/drake-examples/tree/master/gsp).

## Objective and methods

The goal is to search for factors closely associated with the productivity of states in the USA around the 1970s and 1980s. For the sake of simplicity, we use gross state product as a metric of productivity, and we restrict ourselves to multiple linear regression models with three variables. For each of the 84 possible models, we fit the data and then evaluate the root mean squared prediction error (RMSPE).

$$
\begin{aligned}
\text{RMSPE} = \sqrt{(\text{y} - \widehat{y})^T(y - \widehat{y})}
\end{aligned}
$$
Here, $y$ is the vector of observed gross state products in the data, and $\widehat{y}$ is the vector of predicted gross state products under one of the models. We take the best variables to be the triplet in the model with the lowest RMSPE.

## Data

The `Produc` dataset from the [Ecdat package](https://cran.r-project.org/package=Ecdat) contains data on the Gross State Product from 1970 to 1986. Each row is a single observation on a single state for a single year. The dataset has the following variables as columns. See the references later in this report for more details.

- `gsp`: gross state product.
- `state`: the state.
- `year`: the year.
- `pcap`: private capital stock.
- `hwy`: highway and streets.
- `water`: water and sewer facilities.
- `util`: other public buildings and structures.
- `pc`: public capital.
- `emp`: labor input measured by the employment in non-agricultural payrolls.
- `unemp`: state unemployment rate.

```{r masterdata2}
library(Ecdat)
data(Produc)
head(Produc)
```

## Analysis

First, we load the required packages. `drake` is aware of all the packages you load with `library()` or `require()`.

```{r masterpkgs}
library(biglm) # lightweight models, easier to store than with lm()
library(drake)
library(Ecdat) # econometrics datasets
library(ggplot2)
library(knitr)
library(purrr)
library(tidyverse)
```

Next, we construct our plan. The following code uses `drake`'s special new language for generating plans (learn more [here](#plans)).

```{r gspdsl}
predictors <- setdiff(colnames(Produc), "gsp")

# We will try all combinations of three covariates.
combos <- combn(predictors, 3)

# We have a grid of covariates.
x1 <- rlang::syms(combos[1, ])
x2 <- rlang::syms(combos[2, ])
x3 <- rlang::syms(combos[3, ])

# Requires drake >= 7.0.0 or the development version
# at github.com/ropensci/drake.
# Install with remotes::install_github("ropensci/drake").
plan <- drake_plan(
  model = target(
    biglm(gsp ~ x1 + x2 + x3, data = Ecdat::Produc),
    transform = map(x1 = !!x1, x2 = !!x2, x3 = !!x3)
  ),
  rmspe_i = target(
    get_rmspe(model, Ecdat::Produc),
    transform = map(model)
  ),
  rmspe = target(
    do.call(rbind, rmspe_i),
    transform = combine(rmspe_i)
  ),
  plot = ggsave(filename = file_out("rmspe.pdf"), plot = plot_rmspe(rmspe)),
  report = knit(knitr_in("report.Rmd"), file_out("report.md"), quiet = TRUE)
)

plan
```

We also need to define functions for summaries and plots.

```{r masterget_rmspe}
get_rmspe <- function(model_fit, data){
  y <- data$gsp
  yhat <- as.numeric(predict(model_fit, newdata = data))
  terms <- attr(model_fit$terms, "term.labels")
  data.frame(
    rmspe = sqrt(mean((y - yhat)^2)), # nolint
    X1 = terms[1],
    X2 = terms[2],
    X3 = terms[3]
  )
}

plot_rmspe <- function(rmspe){
  ggplot(rmspe) +
    geom_histogram(aes(x = rmspe), bins = 15)
}
```

We have a [`report.Rmd` file ](https://github.com/wlandau/drake-examples/blob/master/gsp/report.Rmd) to summarize our results at the end.

```{r copyreport}
drake_example("gsp")
file.copy(from = "gsp/report.Rmd", to = ".", overwrite = TRUE)
```

We can inspect the project before we run it.

```{r gspvisgraph}
config <- drake_config(plan)
vis_drake_graph(config)
```

Now, we can run the project.

```{r appmake}
make(plan, verbose = FALSE)
```

## Results

Here are the root mean squared prediction errors of all the models.

```{r masterrmspeplot}
results <- readd(rmspe)
library(ggplot2)
plot_rmspe(rmspe = results)
```

And here are the best models. The best variables are in the top row under `X1`, `X2`, and `X3`.

```{r masterbestmodels}
head(results[order(results$rmspe, decreasing = FALSE), ])
```

## Comparison with GNU Make

If we were using [Make](https://www.gnu.org/software/make/) instead of `drake` with the same set of targets, the analogous [Makefile](https://www.gnu.org/software/make/) would look something like this pseudo-code sketch.

<pre><code>models = model_state_year_pcap.rds model_state_year_hwy.rds ... # 84 of these

model_%
    Rscript -e 'saveRDS(lm(...), ...)'

rmspe_%: model_%
    Rscript -e 'saveRDS(get_rmspe(...), ...)'

rmspe.rds: rmspe_%
    Rscript -e 'saveRDS(rbind(...), ...)'

rmspe.pdf: rmspe.rds
    Rscript -e 'ggplot2::ggsave(plot_rmspe(readRDS("rmspe.rds")), "rmspe.pdf")'

report.md: report.Rmd
    Rscript -e 'knitr::knit("report.Rmd")'
</code></pre>

There are three main disadvantages to this approach.

1. Every target requires a new call to `Rscript`, which means that more time is spent initializing R sessions than doing the actual work.
2. The user must micromanage nearly one hundred output files (in this case, `*.rds` files), which is cumbersome, messy, and inconvenient. `drake`, on the other hand, automatically manages storage using a [storr cache](https://github.com/richfitz/storr).
3. The user needs to write the names of the 84 `models` near the top of the `Makefile`, which is less convenient than maintaining a data frame in R.

## References 

- Baltagi, Badi H (2003). Econometric analysis of panel data, John Wiley and sons, http://www.wiley.com/legacy/wileychi/baltagi/.
- Baltagi, B. H. and N. Pinnoi (1995). "Public capital stock and state productivity growth: further evidence", Empirical Economics, 20, 351-359.
- Munnell, A. (1990). "Why has productivity growth declined? Productivity and public investment"", New England Economic Review, 3-22.
- Yves Croissant (2016). Ecdat: Data Sets for Econometrics. R package version 0.3-1. https://CRAN.R-project.org/package=Ecdat.

```{r rmfiles_main03, echo = FALSE}
clean(destroy = TRUE, verbose = FALSE)
unlink(
  c(
    "Makefile", "report.Rmd", "figure", "gsp",
    "shell.sh", "STDIN.o*", "Thumbs.db"
  ),
  recursive = TRUE
)
```
