# Memory management {#memory}

```{r, message = FALSE, warning = FALSE, echo = FALSE}
knitr::opts_knit$set(root.dir = fs::dir_create(tempfile()))
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
library(biglm)
library(drake)
library(tidyverse)
```

The default settings of `drake` prioritize speed over memory efficiency. For projects with large data, this default behavior can cause problems. Consider the following hypothetical workflow, where we simulate several large datasets and summarize them.

```{r, paged.print = FALSE}
reps <- 10 # Serious workflows may have several times more.

# Reduce `n` to lighten the load if you want to try this workflow yourself.
# It is super high in this chapter to motivate the memory issues.
generate_large_data <- function(rep, n = 1e8) {
  tibble(x = rnorm(n), y = rnorm(n), rep = rep)
}

get_means <- function(...) {
  out <- NULL
  for (dataset in list(...)) {
    out <- bind_rows(out, colMeans(dataset))
  }
  out
}

plan <- drake_plan(
  large_data = target(
    generate_large_data(rep),
    transform = map(rep = !!seq_len(reps), .id = FALSE)
  ),
  means = target(
    get_means(large_data),
    transform = combine(large_data)
  ),
  summ = summary(means)
)

print(plan)
```

```{r}
config <- drake_config(plan)
vis_drake_graph(config)
```

If you call `make(plan)` with no additional arguments, `drake` will try to load all the datasets into the same R session. Each dataset from `generate_large_data(n = 1e8)` occupies about 2.4 GB of memory, and most machines cannot handle all the data at once. We should use memory more wisely.

## Garbage collection and custom files

`make()` and `drake_config()` have a `garbage_collection` argument, which tells `drake` to periodically unload data objects that no longer belong to variables. You can also run garbage collection manually with the `gc()` function. For more on garbage collection, please refer to the [memory usage chapter of Advanced R](http://adv-r.had.co.nz/memory.html#gc).

Let's reduce the memory consumption of our example workflow:

1. Call `gc()` after every loop iteration of `get_means()`.
2. Avoid `drake`'s caching system with custom `file_out()` files in the plan.
3. Call `make(plan, garbage_collection = TRUE)`.

```{r, paged.print = FALSE}
reps <- 10 # Serious workflows may have several times more.
files <- paste0(seq_len(reps), ".rds")

generate_large_data <- function(file, n = 1e8) {
  out <- tibble(x = rnorm(n), y = rnorm(n)) # a billion rows
  saveRDS(out, file)
}

get_means <- function(files) {
  out <- NULL
  for (file in files) {
    x <- colMeans(readRDS(file))
    out <- bind_rows(out, x)
    gc() # Use the gc() function here to make sure each x gets unloaded.
  }
  out
}

plan <- drake_plan(
  large_data = target(
    generate_large_data(file = file_out(file)),
    transform = map(file = !!files, .id = FALSE)
  ),
  means = get_means(file_in(!!files)),
  summ = summary(means)
)

print(plan)
```

```{r}
config <- drake_config(plan)
vis_drake_graph(config)
```

```{r, eval = FALSE}
make(plan, garbage_collection = TRUE)
```

## Memory strategies

`make()` and `drake_config()` have a `memory_strategy` argument to customize how `drake` loads and unloads targets. With the right memory strategy, you can rely on `drake`'s built-in caching system without having to bother with messy `file_out()` files.

Each memory strategy follows three stages for each target:

1. Initial discard: before building the target, optionally discard some other targets from the R session. The choice of discards depends on the memory strategy. (Note: we do not actually get the memory back until we call `gc()`.)
2. Initial load: before building the target, optionally load any dependencies that are not already in memory.
3. Final discard: optionally discard or keep the return value after the target finishes building. Either way, the return value is still stored in the cache, so you can load it with `loadd()` and `readd()`.

The implementation of these steps varies from strategy to strategy.

Memory strategy | Initial discard | Initial load | Final discard
---|---|---|---
"speed" | Discard nothing | Load any missing dependencies. | Keep the return value loaded.
"autoclean"[^1] | Discard all targets which are not dependencies of the current target. | Load any missing dependencies. | Discard the return value.
"preclean" | Discard all targets which are not dependencies of the current target. | Load any missing dependencies. | Keep the return value loaded.
"lookahead" | Discard all targets which are not dependencies of either (1) the current target or (2) other targets waiting to be checked or built. | Load any missing dependencies. | Keep the return value loaded.
"unload"[^1] | Unload all targets. | Load nothing. | Discard the return value.
"none"[^1] | Unload nothing. | Load nothing. | Discard the return value.

[^1]: Only supported in `drake` version 7.4.0 and above.

With the `"speed"`, `"autoclean"`, `"preclean"`, and `"lookahead"` strategies, you can simply call `make(plan, memory_strategy = YOUR_CHOICE, garbage_collection = TRUE)` and trust that your targets will build normally. For the `"unload"` and `"none"` strategies, there is extra work to do: you will need to manually load each target's dependencies with `loadd()` or `readd()`. This manual bookkeeping lets you aggressively optimize your workflow, and it is less cumbersome than swarms of `file_out()` files. It is particularly useful when you have a large `combine()` step.

Let's redesign the workflow to reap the benefits of `make(plan, memory_strategy = "none", garbage_collection = TRUE)`. The trick is to use [`match.call()`](https://www.rdocumentation.org/packages/base/versions/3.6.0/topics/match.call) inside `get_means()` so we can load and unload dependencies one at a time instead of all at once.

```{r, paged.print = FALSE}
reps <- 10 # Serious workflows may have several times more.

generate_large_data <- function(rep, n = 1e8) {
  tibble(x = rnorm(n), y = rnorm(n), rep = rep)
}

# Load targets one at a time
get_means <- function(...) {
  arg_symbols <- match.call(expand.dots = FALSE)$...
  arg_names <- as.character(arg_symbols)
  out <- NULL
  for (arg_name in arg_names) {
    dataset <- readd(arg_name, character_only = TRUE)
    out <- bind_rows(out, colMeans(dataset))
    gc() # Run garbage collection.
  }
  out
}

plan <- drake_plan(
  large_data = target(
    generate_large_data(rep),
    transform = map(rep = !!seq_len(reps), .id = FALSE)
  ),
  means = target(
    get_means(large_data),
    transform = combine(large_data)
  ),
  summ = {
    loadd(means) # Annoying, but necessary with the "none" strategy.
    summary(means)
  }
)
```

Now, we can build our targets.

```{r, eval = FALSE}
make(plan, memory_strategy = "none", garbage_collection = TRUE)
```

But there is a snag: we needed to manually load `means` in the command for `summ` (notice the call to `loadd()`). This is annoying, especially because `means` is quite small. Fortunately, `drake` lets you define different memory strategies for different targets in the plan. The target-specific memory strategies override the global one (i.e. the `memory_strategy` argument of `make()` and `drake_config()`).

```{r}
plan <- drake_plan(
  large_data = target(
    generate_large_data(rep),
    transform = map(rep = !!seq_len(reps), .id = FALSE),
    memory_strategy = "none"
  ),
  means = target(
    get_means(large_data),
    transform = combine(large_data),
    memory_strategy = "unload" # Be careful with this one.
  ),
  summ = summary(means)
)

print(plan)
```

In fact, now you can run `make()` without setting a global memory strategy at all.

```{r, eval = FALSE}
make(plan, garbage_collection = TRUE)
```

## Data splitting

The [`split()` transformation](https://ropenscilabs.github.io/drake-manual/plans.html#split) breaks up a dataset into smaller targets. The ordinary use of `split()` is to partition an in-memory dataset into slices.

```{r, paged.print = FALSE}
drake_plan(
  data = get_large_data(),
  x = target(
    data %>%
      analyze_data(),
    transform = split(data, slices = 4)
  )
)
```

However, you can also use it to load individual pieces of a large file, thus conserving memory. The trick is to break up an index set instead of the data itself. In the following sketch, `get_number_of_rows()` and `read_selected_rows()` are user-defined functions, and `%>%` is the [`magrittr`](https://magrittr.tidyverse.org) pipe.

```{r, paged.print = FALSE}
get_number_of_rows <- function(file) {
  # ...
}

read_selected_rows <- function(which_rows, file) {
  # ...
}

plan <- drake_plan(
  row_indices = file_in("large_file.csv") %>%
    get_number_of_rows() %>%
    seq_len(),
  subset = target(
    row_indices %>%
      read_selected_rows(file = file_in("large_file.csv")),
    transform = split(row_indices, slices = 4)
  )
)

plan

drake_plan_source(plan)
```
