# The mtcars example and `drake` plan generation {#mtcars}

```{r suppression04, echo = F}
suppressMessages(suppressWarnings(library(drake)))
clean(destroy = TRUE, verbose = FALSE)
unlink(
  c("mtcars", "Makefile", "report.Rmd", "shell.sh", "STDIN.o*", "Thumbs.db"),
  recursive = TRUE
)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This chapter is a walkthrough of `drake`'s main functionality based on the mtcars example. It sets up the project and runs it repeatedly to demonstrate `drake`'s most important functionality.

## Get the code.

Write the code files to your workspace.

```{r getthequickstartcode04, eval = FALSE}
drake_example("mtcars")
```

The new `mtcars` folder now includes a file structure of a serious `drake` project, plus an `interactive-tutorial.R` to narrate the example. The code is also [online here](https://github.com/wlandau/drake-examples/tree/master/mtcars).

## Quick examples

Inspect and run your project.

```{r quickstartquickstart, eval = FALSE}
library(drake)
load_mtcars_example()            # Get the code with drake_example("mtcars").
config <- drake_config(my_plan) # Master configuration list
vis_drake_graph(config)         # Hover, click, drag, zoom, pan.
make(my_plan)                   # Run the workflow.
outdated(config)                # Everything is up to date.
```

Debug errors.

```{r quickdebug, eval = FALSE}
failed()                   # Targets that failed in the most recent `make()`
context <- diagnose(large) # Diagnostic metadata: errors, warnings, etc.
error <- context$error
str(error)                 # Object of class "error"
error$message
error$call
error$calls                # Full traceback of nested calls leading up to the error. # nolint
```

Dive deeper into the built-in examples.

```{r noeval2, eval = FALSE}
drake_example("mtcars") # Write the code files.
drake_examples()        # List the other examples.
```

## The motivation of the mtcars example

Is there an association between the weight and the fuel efficiency of cars? To find out, we use the `mtcars` dataset from the `datasets` package. The `mtcars` dataset originally came from the 1974 Motor Trend US magazine, and it contains design and performance data on 32 models of automobile.

```{r mtcarsquickstart}
# ?mtcars # more info
head(mtcars)
```

Here, `wt` is weight in tons, and `mpg` is fuel efficiency in miles per gallon. We want to figure out if there is an association between `wt` and `mpg`. The `mtcars` dataset itself only has 32 rows, so we generate two larger bootstrapped datasets and then analyze them with regression models. We summarize the regression models to see if there is an association.

## Set up the mtcars example

Before you run your project, you need to set up the workspace. In other words, you need to gather the "imports": functions, pre-loaded data objects, and saved files that you want to be available before the real work begins.

```{r libs}
library(knitr) # drake knows which packages you load.
library(drake)
```

We need a function to bootstrap larger datasets from `mtcars`.

```{r sim}
# Pick a random subset of n rows from a dataset
random_rows <- function(data, n){
  data[sample.int(n = nrow(data), size = n, replace = TRUE), ]
}

# Bootstrapped datasets from mtcars.
simulate <- function(n){
  # Pick a random set of cars to bootstrap from the mtcars data.
  data <- random_rows(data = mtcars, n = n)

  # x is the car's weight, and y is the fuel efficiency.
  data.frame(
    x = data$wt,
    y = data$mpg
  )
}
```

We also need functions to apply the regression models we need for detecting associations.

```{r reg}
# Is fuel efficiency linearly related to weight?
reg1 <- function(d){
  lm(y ~ + x, data = d)
}

# Is fuel efficiency related to the SQUARE of the weight?
reg2 <- function(d){
  d$x2 <- d$x ^ 2
  lm(y ~ x2, data = d)
}
```

We want to summarize the final results in an R Markdown report, so we need the the `report.Rmd` source file. You can get it with `drake_example("mtcars")` or `load_mtcars_example()`.

```{r file}
drake_example("mtcars", overwrite = TRUE)
file.copy("mtcars/report.Rmd", ".", overwrite = TRUE)
```

Here are the contents of the report. It will serve as a final summary of our work, and we will process it at the very end. Admittedly, some of the text spoils the punch line.

```{r readlinesofreport}
cat(readLines("report.Rmd"), sep = "\n")
```

Now, all our imports are set up. When the real work begins, `drake` will import functions and data objects from your R session environment

```{r robjimportsquickstart}
ls()
```

and saved files from your file system.

```{r filesystemimportsquickstart}
list.files()
```

## The `drake` plan

Now that your workspace of imports is prepared, we can outline the real work step by step in a `drake` plan.

```{r previewmyplan}
load_mtcars_example() # Get the code with drake_example("mtcars").
my_plan
```

Each row is an intermediate step, and each **command** generates a single **target**. A target is an output R object (cached when generated) or an output file (specified with single quotes), and a command just an ordinary piece of R code (not necessarily a single function call). Commands make use of targets generated by other commands, objects your environment, input files, and namespaced objects/functions from packages (referenced with `::` or `:::`). These dependencies give your project an underlying network representation.

```{r graph1quick}
# Hover, click, drag, zoom, and pan.
config <- drake_config(my_plan)
vis_drake_graph(config, width = "100%", height = "500px") # Also drake_graph()
```

You can also check the dependencies of individual targets and imported functions.

```{r checkdeps}
deps_code(reg2)

deps_code(my_plan$command[1]) # Files like report.Rmd are single-quoted.

deps_code(my_plan$command[nrow(my_plan)])
```

List all the reproducibly-tracked objects and files.

```{r tracked}
tracked(config)
```

## Generate the plan

### The easy way

`drake` version 7.0.0 will support new special syntax to create complicated `drake` plans from boilerplate code. See the [chapter on plans](#plans) for more details. To get the funcionality early, install development drake.

```{r installdevdrakedslmtcars, eval = FALSE}
install.packages("remotes")
library(remotes)
install_github("ropensci/drake")
```

Then, use transformations to generate the plan.

```{r, dslmtcars}
drake_plan(
  report = knit(knitr_in("report.Rmd"), file_out("report.md"), quiet = TRUE),
  small = simulate(48),
  large = simulate(64),
  regression1 = target(
    reg1(data),
    transform = map(data = c(small, large)),
    group = reg
  ),
  regression2 = target(
    reg2(data),
    transform = map(data),
    group = reg
  ),
  summ = target(
    suppressWarnings(summary(reg$residuals)),
    transform = map(reg)
  ),
  coef = target(
    suppressWarnings(summary(reg))$coefficients,
    transform = map(reg)
  )
)
```

In the first row above, `knitr_in()` indicates that `report.Rmd` is a dependency and targets loaded with `loadd()` and `readd()` in active code chunks are also dependencies. Use `file_out()` to tell `drake` that the target is a file output.

### The old way

`drake` has old wildcard templating functions to help generate plans. It is more difficult to adapt them to practical use cases, but they have been around since the early days of `drake`.

Here are the commands to generate the bootstrapped datasets.

```{r datasets}
my_datasets <- drake_plan(
  small = simulate(48),
  large = simulate(64))
my_datasets
```

For multiple replicates:

```{r expand}
expand_plan(my_datasets, values = c("rep1", "rep2"))
```

Here is a template for applying our regression models to our bootstrapped datasets.

```{r methods}
methods <- drake_plan(
  regression1 = reg1(dataset__),
  regression2 = reg2(dataset__))
methods
```

We evaluate the `dataset__` wildcard to generate all the regression commands we need.

```{r analyses}
my_analyses <- evaluate_plan(
  methods, wildcard = "dataset__",
  values = my_datasets$target
)
my_analyses
```

Next, we summarize each analysis of each dataset. We calculate descriptive statistics on the residuals, and we collect the regression coefficients and their p-values.

```{r summaries}
summary_types <- drake_plan(
  summ = suppressWarnings(summary(analysis__$residuals)),
  coef = suppressWarnings(summary(analysis__))$coefficients
)
summary_types

my_summaries <- evaluate_plan(
  summary_types,
  wildcard = "analysis__",
  values = my_analyses$target
)
my_summaries
```

For your `knitr` reports, use `knitr_in()` in your commands so that `report.Rmd` is a dependency and targets loaded with `loadd()` and `readd()` in active code chunks are also dependencies. Use `file_out()` to tell `drake` that the target is a file output.

```{r reportplan}
report <- drake_plan(
  report = knit(knitr_in("report.Rmd"), file_out("report.md"), quiet = TRUE)
)
report
```

Finally, consolidate your workflow using `rbind()`. Row order does not matter.

```{r wholeplan04}
my_plan <- rbind(report, my_datasets, my_analyses, my_summaries)
my_plan
```

## Run the workflow

You may want to check for outdated or missing targets/imports first.

```{r firstmake}
config <- drake_config(my_plan, verbose = FALSE)
outdated(config) # Targets that need to be (re)built.

missed(config) # Checks your workspace.
```

Then just `make(my_plan)`.

```{r firstmakeforreal}
make(my_plan)
```

For the `reg2()` model on the small dataset, the p-value on `x2` is so small that there may be an association between weight and fuel efficiency after all.

```{r getmtcarsanswer}
readd(coef_regression2_small)
```

The non-file dependencies of your last target are already loaded in your workspace.

```{r autoload}
ls()
```

```{r plotgraphfirstmake}
outdated(config) # Everything is up to date.

build_times(digits = 4) # How long did it take to make each target?
```

See also `predict_runtime()` and `rate_limiting_times()`.

In the new graph, the black nodes from before are now green.

```{r graph2quick}
# Hover, click, drag, zoom, and explore.
vis_drake_graph(config, width = "100%", height = "500px")
```

Optionally, get [visNetwork](http://datastorm-open.github.io/visNetwork/) nodes and edges so you can make your own plot with `visNetwork()` or `render_drake_graph()`.

```{r dfgraph2quick, eval = FALSE}
drake_graph_info(config)
```

Use `readd()` and `loadd()` to load targets into your workspace. (They are cached in the hidden `.drake/` folder using [storr](https://CRAN.R-project.org/package=storr)). There are many more functions for interacting with the cache.

```{r cache}
readd(coef_regression2_large)

loadd(small)

head(small)

rm(small)

cached()

progress()

# drake_session() # sessionInfo() of the last make() # nolint
```

The next time you run `make(my_plan)`, nothing will build because `drake` knows everything is already up to date.

```{r uptodateinvig}
make(my_plan)
```

But if you change one of your functions, commands, or other dependencies, drake will update the affected targets. Suppose we change the quadratic term to a cubic term in `reg2()`. We might want to do this if we suspect a cubic relationship between tons and miles per gallon.

```{r changereg2inmtcarscase}
reg2 <- function(d) {
  d$x3 <- d$x ^ 3
  lm(y ~ x3, data = d)
}
```

The targets that depend on `reg2()` need to be rebuilt.

```{r plotwithreg2}
config <- drake_config(my_plan)
outdated(config)
```

**Advanced**: To get a rough idea of why a target is out of date, you can use `dependency_profile()`. It will tell you if any of the following changed since the last `make()`:

- The command in the `drake` plan.
- At least one non-file dependency. (For this, the imports have to be up to date and cached, either with `make()`, `make(skip_targets = TRUE)`, `outdated()`, or similar.)
- At least one input file declared with `file_in()` or `knitr_in()`.
- At least one output file declared with `file_out()`.

```{r depprofile}
dependency_profile(target = regression2_small, config = config)
```

```{r graph3quick}
# Hover, click, drag, zoom, and explore.
vis_drake_graph(config, width = "100%", height = "500px")
```

The next `make()` will rebuild the targets depending on `reg2()` and leave everything else alone.

```{r remakewithreg2}
make(my_plan)
```

Trivial changes to whitespace and comments are totally ignored.

```{r trivial}
reg2 <- function(d) {
  d$x3 <- d$x ^ 3
    lm(y ~ x3, data = d) # I indented here.
}
outdated(config) # Everything is up to date.
```

drake cares about nested functions too: nontrivial changes to `random_rows()` will propagate to `simulate()`  and all the downstream targets.

```{r, changerandomrows}
random_rows <- function(data, n){
  n <- n + 1
  data[sample.int(n = nrow(data), size = n, replace = TRUE), ]
}

outdated(config)

make(my_plan)
```

Need to add new work on the fly? Just append rows to the `drake` plan. If the rest of your workflow is up to date, only the new work is run.

```{r newstuff}
new_simulation <- function(n){
  data.frame(x = rnorm(n), y = rnorm(n))
}

additions <- drake_plan(
  new_data = new_simulation(36) + sqrt(10))
additions

my_plan <- rbind(my_plan, additions)
my_plan

make(my_plan)
```

If you ever need to erase your work, use `clean()`. The next `make()` will rebuild any cleaned targets, so be careful. You may notice that by default, the size of the cache does not go down very much. To purge old data, you could use `clean(garbage_collection = TRUE, purge = TRUE)`. To do garbage collection without removing any important targets, use `drake_gc()`.

```{r cleanup}
# Uncaches individual targets and imported objects.
clean(small, reg1, verbose = FALSE)
clean(verbose = FALSE) # Cleans all targets out of the cache.
drake_gc(verbose = FALSE) # Just garbage collection.
clean(destroy = TRUE, verbose = FALSE) # removes the cache entirely
```

```{r endofline_quickstart04, echo = F}
clean(destroy = TRUE, verbose = FALSE)
unlink(
  c("mtcars", "Makefile", "report.Rmd", "shell.sh", "STDIN.o*", "Thumbs.db"),
  recursive = TRUE
)
```
