# `drake` projects {#projects}

```{r, message = FALSE, warning = FALSE, echo = FALSE}
knitr::opts_knit$set(root.dir = fs::dir_create(tempfile()))
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(
  drake_make_menu = FALSE,
  drake_clean_menu = FALSE,
  warnPartialMatchArgs = FALSE,
  crayon.enabled = FALSE,
  readr.show_progress = FALSE,
  tidyverse.quiet = TRUE
)
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
library(drake)
library(dplyr)
library(ggplot2)
drake_example("main")
tmp <- file.copy("main/R", ".", recursive = TRUE)
tmp <- file.copy("main/_drake.R", ".")
tmp <- file.copy("main/raw_data.xlsx", ".")
tmp <- file.copy("main/", ".")
tmp <- file.copy("main/report.Rmd", ".")
tmp <- file.copy("main/report.Rmd", ".")
tmp <- file.copy("main/report.Rmd", ".")

```

`drake`'s design philosophy is extremely R-focused. It embraces in-memory configuration, in-memory dependencies, interactivity, and flexibility. This scope leaves project setup and file management decisions mostly up to the user. This chapter tries to fill in the blanks and address practical hurdles when it comes to setting up projects.

## External resources

- [Miles McBain](https://github.com/MilesMcBain)'s [excellent blog post](https://milesmcbain.xyz/the-drake-post/) explains the practical issues {drake} solves for most projects, how to set up a project as quickly and painlessly as possible, and how to  overcome common obstacles.
- Miles' [`dflow`](https://github.com/MilesMcBain/dflow) package generates the file structure for a boilerplate `drake` project. It is a more thorough alternative to `drake::use_drake()`.
- `drake` is heavily function-oriented by design, and Miles' [`fnmate`](https://github.com/MilesMcBain/fnmate) package automatically generates boilerplate code and docstrings for functions you mention in `drake` plans.

## Code files

The names and locations of the files are entirely up to you, but this pattern is particularly useful to start with.

```
make.R
R/
├── packages.R
├── functions.R
└── plan.R
```

Here, `make.R` is a master script that

1. Loads your packages, functions, and other in-memory data.
2. Creates the `drake` plan.
3. Calls `make()`.

Let's consider the [main example](https://github.com/wlandau/drake-examples/tree/master/main), which you can download with `drake_example("main")`. Here, our master script is called `make.R`:

```{r, eval = FALSE}
source("R/packages.R")  # loads packages
source("R/functions.R") # defines the create_plot() function
source("R/plan.R")      # creates the drake plan
# options(clustermq.scheduler = "multicore") # optional parallel computing. Also needs parallelism = "clustermq"
make(
  plan, # defined in R/plan.R
  verbose = 2
)
```

We have an `R` folder containing our supporting files. `packages.R` typically includes all the packages you will use in the workflow. 

```{r, eval = FALSE}
# packages.R
library(drake)
library(dplyr)
library(ggplot2)
```

Your `functions.R` typically has the supporting custom functions you write for the workflow. If there are many functions, you could split them up into multiple files.

```{r, eval = FALSE}
# functions.R
create_plot <- function(data) {
  ggplot(data) +
    geom_histogram(aes(x = Ozone), binwidth = 10) +
    theme_gray(24)
}
```

Finally, it is good practice to define a `plan.R` that defines the plan.

```{r, eval = FALSE}
# plan.R
plan <- drake_plan(
  raw_data = readxl::read_excel(file_in("raw_data.xlsx")),
  data = raw_data %>%
    mutate(Ozone = replace_na(Ozone, mean(Ozone, na.rm = TRUE))),
  hist = create_plot(data),
  fit = lm(Ozone ~ Wind + Temp, data),
  report = rmarkdown::render(
    knitr_in("report.Rmd"),
    output_file = file_out("report.html"),
    quiet = TRUE
  )
)
```

To run the example project above,

1. Start a clean new R session.
2. Run the `make.R` script.

On Mac and Linux, you can do this by opening a terminal and entering `R CMD BATCH --no-save make.R`. On Windows, restart your R session and call `source("make.R")` in the R console.

Note: this part of `drake` does not inherently focus on your script files. There is nothing magical about the names `make.R`, `packages.R`, `functions.R`, or `plan.R`. Different projects may require different file structures.

`drake` has other functions to inspect your results and examine your workflow. Before invoking them interactively, it is best to start with a clean new R session.

```{r, eval = FALSE}
# Restart R.
interactive()
#> [1] TRUE
source("R/packages.R")
source("R/functions.R")
source("R/plan.R")
vis_drake_graph(plan)
```

## Safer interactivity

### Motivation

A serious [`drake`](https://github.com/ropensci/drake) workflow should be consistent and reliable, ideally with the help of a [master R script](https://github.com/wlandau/drake-examples/blob/master/gsp/make.R). Before it builds your targets, this script should begin in a fresh R session and load your packages and functions in a [dependable manner](https://github.com/wlandau/drake-examples/blob/d9417547a05aec416afbbda913eaf2d44a552d5b/gsp/make.R#L4-L6). [Batch mode](https://www.statmethods.net/interface/batch.html) makes sure all this goes according to plan.

If you use a single persistent [interactive R session](https://stat.ethz.ch/R-manual/R-devel/library/base/html/interactive.html) to repeatedly invoke `make()` while you develop the workflow, then over time, your session could grow stale and accidentally invalidate targets. For example, if you interactively tinker with a new version of `create_plot()`, targets `hist` and `report` will fall out of date without warning, and the next `make()` will build them again. Even worse, the outputs from `hist` and `report` will be wrong if they depend on a half-finished `create_plot()`.

The quickest workaround is to restart R and `source()` your setup scripts all over again. However, a better solution is to use [`r_make()`](https://docs.ropensci.org/drake/reference/r_make.html) and friends. [`r_make()`](https://docs.ropensci.org/drake/reference/r_make.html) runs `make()` in a new transient R session so that accidental changes to your interactive environment do not break your workflow.

### Usage

To use [`r_make()`](https://docs.ropensci.org/drake/reference/r_make.html), you need a configuration R script. Unless you supply a custom file path (e.g. `r_make(source = "your_file.R")` or `options(drake_source = "your_file.R")`) `drake` assumes this configuration script is called `_drake.R`. (So the file name really *is* magical in this case). The suggested file structure becomes:

```
_drake.R
R/
├── packages.R
├── functions.R
└── plan.R
```

Like our previous `make.R` script, `_drake.R` runs all our pre-`make()` setup steps. But this time, rather than calling `make()`, it ends with a call to `drake_config()`. `drake_config()` is the initial preprocessing stage of `make()`, and it accepts all the same arguments as `make()`.

Example `_drake.R`:

```{r, eval = FALSE}
source("R/packages.R")
source("R/functions.R")
source("R/plan.R")
# options(clustermq.scheduler = "multicore") # optional parallel computing
drake_config(plan, verbose = 2)
```

Here is what happens when you call `r_make()`.

1. `drake` launches a new transient R session using [`callr::r()`](https://github.com/r-lib/callr). The remaining steps all happen within this transient session.
2. Run the configuration script (e.g. `_drake.R`) to
    a. Load the packages, functions, global options, `drake` plan, etc. into the session's environnment, and
    b. Run the call to `drake_config()`and store the results in a variable called `config`.
3. Execute `make_impl(config = config)`, an internal `drake` function.

The purpose of `drake_config()` is to collect and sanitize all the parameters and settings that `make()` needs to do its job. In fact, if you do not set the `config` argument explicitly, then `make()` invokes `drake_config()` behind the scenes. `make(plan, parallelism = "clustermq", jobs = 2, verbose = 6)` is equivalent to

```{r, eval = FALSE}
config <- drake_config(plan, verbose = 2)
make_impl(config = config)
```

There are many more `r_*()` functions besides `r_make()`, each of which launches a fresh session and runs an inner `drake` function on the `config` object from `_drake.R`.

Outer function call | Inner function call
--- | ---
`r_make()` | `make_impl(config = config)`
`r_drake_build(...)` | `drake_build_impl(config, ...)`
`r_outdated(...)` | `outdated_impl(config, ...)`
`r_missed(...)` | `missed_impl(config, ...)`
`r_vis_drake_graph(...)` | `vis_drake_graph_impl(config, ...)`
`r_sankey_drake_graph(...)` | `sankey_drake_graph_impl(config, ...)`
`r_drake_ggraph(...)` | `drake_ggraph_impl(config, ...)`
`r_drake_graph_info(...)` | `drake_graph_info_impl(config, ...)`
`r_predict_runtime(...)` | `predict_runtime_impl(config, ...)`
`r_predict_workers(...)` | `predict_workers_impl(config, ...)`


```{r}
clean()
r_outdated(r_args = list(show = FALSE))

r_make()
r_outdated(r_args = list(show = FALSE))

r_vis_drake_graph(targets_only = TRUE, r_args = list(show = FALSE))
```

Remarks:

- You can run `r_make()` in an interactive session, but the transient process it launches will not be interactive. Thus, any `browser()` statements in the commands in your `drake` plan will be ignored.
- You can select and configure the underlying [`callr`](https://github.com/r-lib/callr) function using arguments `r_fn` and `r_args`, respectively.
- For example code, you can download the updated [main example](https://github.com/wlandau/drake-examples/tree/master/main) (`drake_example("main")`) and experiment with files [`_drake.R`](https://github.com/wlandau/drake-examples/blob/master/main/_drake.R) and [`interactive.R`](https://github.com/wlandau/drake-examples/blob/master/main/interactive.R).

## Script file pitfalls

Despite the above discussion of R scripts, `drake` plans rely more on in-memory functions. You might be tempted to write a plan like the following, but then `drake` cannot tell that `my_analysis` depends on `my_data`.

```{r}
bad_plan <- drake_plan(
  my_data = source(file_in("get_data.R")),
  my_analysis = source(file_in("analyze_data.R")),
  my_summaries = source(file_in("summarize_data.R"))
)

vis_drake_graph(bad_plan, targets_only = TRUE)
```

When it comes to plans, use *functions* instead.

```{r, eval = FALSE}
source("my_functions.R") # defines get_data(), analyze_data(), etc.
good_plan <- drake_plan(
  my_data = get_data(file_in("data.csv")), # External files need to be in commands explicitly. # nolint
  my_analysis = analyze_data(my_data),
  my_summaries = summarize_results(my_data, my_analysis)
)

vis_drake_graph(good_plan, targets_only = TRUE)
```

In `drake` >= 7.6.2.9000, code_to_function() leverages existing imperative scripts for use in a `drake` plan.

```{r, eval=FALSE}
get_data <- code_to_function("get_data.R")
do_analysis <- code_to_function("analyze_data.R")
do_summary <- code_to_function("summarize_data.R")

good_plan <- drake_plan(
  my_data = get_data(),
  my_analysis = do_analysis(my_data),
  my_summaries = do_summary(my_data, my_analysis)
)

vis_drake_graph(good_plan, targets_only = TRUE)

```

```{r, echo = FALSE}
good_plan <- drake_plan(
  my_data = get_data(file_in("data.csv")), # External files need to be in commands explicitly. # nolint
  my_analysis = analyze_data(my_data),
  my_summaries = summarize_results(my_data, my_analysis)
)

vis_drake_graph(good_plan, targets_only = TRUE)
```


## Workflows as R packages

The R package structure is a great way to organize and quality-control a data analysis project. If you write a `drake` workflow as a package, you will need

1. Supply the namespace of your package to the `envir` argument of `make()` or `drake_config()` (e.g. `make(envir = getNamespace("yourPackage")` so `drake` can watch you package's functions for changes and rebuild downstream targets accordingly.
2. If you load the package with `devtools::load_all()`, set the `prework` argument of `make()`: e.g. `make(prework = "devtools::load_all()")` and custom set the `packages` argument so your package name is not included. (Everything in `packages` is loaded with `library()`).

For a minimal example, see [Tiernan Martin](https://github.com/tiernanmartin)'s [`drakepkg`](https://github.com/tiernanmartin/drakepkg).

## Other tools

[`drake`](https://github.com/ropensci/drake) enhances reproducibility, but not in all respects. [Local library managers](https://rstudio.github.io/packrat), [containerization](https://www.docker.com), and [session management tools](https://github.com/tidyverse/reprex) offer more robust solutions in their respective domains. Reproducibility encompasses a [wide variety of tools and techniques](https://github.com/karthik/rstudio2019) all working together. Comprehensive overviews:

- [PLOS article](https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1005510) by Wilson et al.
- [RStudio Conference 2019 presentation ](https://github.com/karthik/rstudio2019) by [Karthik Ram](https://github.com/karthik).
- [`rrtools`](https://github.com/benmarwick/rrtools) by [Ben Marwick](https://github.com/benmarwick).
