# Storage {#storage}

```{r, message = FALSE, warning = FALSE, echo = FALSE}
knitr::opts_knit$set(root.dir = fs::dir_create(tempfile()))
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(
  drake_make_menu = FALSE,
  drake_clean_menu = FALSE,
  warnPartialMatchArgs = FALSE,
  crayon.enabled = FALSE,
  readr.show_progress = FALSE,
  tidyverse.quiet = TRUE
)
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
library(drake)
library(tidyverse)
```

## `drake`'s cache

When you run `make()`, `drake` stores your targets in a hidden storage cache.

```{r}
library(drake)
load_mtcars_example() # from https://github.com/wlandau/drake-examples/tree/main/mtcars
make(my_plan, verbose = 0L)
```

The default cache is a hidden `.drake` folder. 

```{r, eval = FALSE}
find_cache()
### [1] "/home/you/project/.drake"
```

`drake`'s `loadd()` and `readd()` functions load targets into memory.

```{r}
loadd(large)
head(large)

head(readd(small))
```


## Efficient target storage

`drake` supports custom formats for large and specialized targets. For example, the `"fst"` format uses the [`fst`](https://github.com/fstpackage/fst) package to save data frames faster. Simply enclose the command and the format together with the `target()` function.

```{r, eval = FALSE}
library(drake)
n <- 1e8 # Each target is 1.6 GB in memory.
plan <- drake_plan(
  data_fst = target(
    data.frame(x = runif(n), y = runif(n)),
    format = "fst"
  ),
  data_old = data.frame(x = runif(n), y = runif(n))
)
make(plan)
#> target data_fst
#> target data_old
build_times(type = "build")
#> # A tibble: 2 x 4
#>   target   elapsed              user                 system    
#>   <chr>    <Duration>           <Duration>           <Duration>
#> 1 data_fst 13.93s               37.562s              7.954s    
#> 2 data_old 184s (~3.07 minutes) 177s (~2.95 minutes) 4.157s
```

For more details and a complete list of formats, see <https://books.ropensci.org/drake/plans.html#special-data-formats-for-targets>.


## Why is my cache so big?

### Old targets

By default, `drake` holds on to all your targets from all your runs of `make()`. Even if you run `clean()`, the data stays in the cache in case you need to recover it.

```{r}
clean()

make(my_plan, recover = TRUE)
```

If you really want to remove old historical values of targets, run `drake_gc()` or  `drake_cache()$gc()`.

```{r}
drake_gc()
```

`clean()` also has a `garbage_collection` argument for this purpose. Here is a slick way to remove historical targets and targets no longer in your plan.

```{r}
clean(list = cached_unplanned(my_plan), garbage_collection = TRUE)
```

### Garbage from interrupted builds

If `make()` crashes or gets interrupted, old files can accumulate in `.drake/scratch/` and `.drake/drake/tmp/`. As long as `make()` is no longer running, can safely remove the files in those folders (but keep the folders themselves).

## Interfaces to the cache

`drake` uses the [storr](https://github.com/richfitz/storr) package to create and modify caches.

```{r}
library(storr)
cache <- storr_rds(".drake")

head(cache$list())

head(cache$get("small"))
```

`drake` has its own interface on top of [storr](https://github.com/richfitz/storr) to make it easier to work with the default  `.drake/` cache. The `loadd()`, `readd()`, and `cached()` functions explore saved targets.

```{r}
head(cached())

head(readd(small))

loadd(large)

head(large)

rm(large) # Does not remove `large` from the cache.
```

`new_cache()` create caches and `drake_cache()` recovers existing ones. (`drake_cache()` is only supported in `drake` version 7.4.0 and above.)

```{r}
cache <- drake_cache()
cache$driver$path

cache <- drake_cache(path = ".drake") # File path to drake's cache.
cache$driver$path
```

You can supply your own cache to `make()` and friends (including specialized `storr` caches like [`storr_dbi()`](http://richfitz.github.io/storr/reference/storr_dbi.html)).

```{r}
plan <- drake_plan(x = 1, y = sqrt(x))
make(plan, cache = cache)

vis_drake_graph(plan, cache = cache)
```

Destroy caches to remove them from your file system.

```{r}
cache$destroy()

file.exists(".drake")
```
