# Triggers: decision rules for building targets {#triggers}

```{r, message = FALSE, warning = FALSE, include = FALSE}
knitr::opts_knit$set(root.dir = fs::dir_create(tempfile()))
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
options(
  drake_make_menu = FALSE,
  drake_clean_menu = FALSE,
  warnPartialMatchArgs = FALSE,
  crayon.enabled = FALSE,
  readr.show_progress = FALSE,
  tidyverse.quiet = TRUE
)
```

```{r, message = FALSE, warning = FALSE, include = FALSE}
library(drake)
library(tidyverse)
invisible(drake_example("main", overwrite = TRUE))
invisible(file.copy("main/raw_data.xlsx", ".", overwrite = TRUE))
invisible(file.copy("main/report.Rmd", ".", overwrite = TRUE))
```

When you call `make()`, `drake` tries to skip as many targets as possible. If it thinks a command will return the same value as last time, it does not bother running it. In other words, `drake` is lazy, and laziness saves you time.

## What are triggers?

To figure out whether it can skip a target, `drake` goes through an intricate checklist of **triggers**:

1. The **missing** trigger: Do we lack a return value from a previous `make()`? Maybe you are building the target for the first time or you removed it from the cache with `clean()`.
2. The **command** trigger: did the command in the `drake` plan change nontrivially since the last `make()`? Changes to spacing, formatting, and comments are ignored.
3. The **depend** trigger: did any non-file dependencies change since the last `make()`? These could be:
    - Other targets.
    - Imported objects.
    - Imported functions. To track changes to a function, `drake` removes any code closed in `ignore()`, deparses the literal code so that whitespace is standardized and comments are removed, and then hashes the resulting string. In some cases, `drake` makes special adjustments for strange edge cases like [`Rcpp` functions with pointers](https://github.com/ropensci/drake/issues/806) and functions defined with `Vectorize()`. However, edge cases like [this one](https://github.com/ropensci-books/drake/issues/130#issue-522436582) are inevitable because of the flexibility of R.
    - Any dependencies of imported functions.
    - Any dependencies of dependencies of imported functions, and so on.
4. The **file** trigger: did any file inputs or file outputs change since the last `make()`? These files are the ones explicitly declared in the command with `file_in()`, `knitr_in()`, and `file_out()`.
5. The **seed** trigger: for statistical reproducibility, `drake` assigns a unique seed to each target based on the target's name and the global `seed` argument to `make()`. If you change the target's pseudo-random number generator seed either with the `seed` argument or the custom `seed` column in the plan, this change will cause a rebuild if the `seed` trigger is turned on.
6. The **format** trigger: did you add or change the target's storage format since last build? Details: <https://books.ropensci.org/drake/plans.html#special-data-formats-for-targets>.
7. The **condition** trigger: an optional user-defined piece of code that evaluates to a `TRUE`/`FALSE` value. The target builds if the value is `TRUE`.
8. The **change** trigger:  an optional user-defined piece of code that evaluates to any value (preferably small and quick to compute). The target builds if the value changed since the last `make()`.

If *any* trigger detects something wrong or different with the target or its dependencies, the next `make()` will run the command and (re)build the target.

## Customization

With the `trigger()` function, you can create your own customized checklist of triggers. Let's run a simple workflow with just the **missing** trigger. We deactivate the **command**, **depend**, and **file** triggers by setting the respective `command`, `depend`, and `file` arguments to `FALSE`.

```{r}
plan <- drake_plan(
  psi_1 = (sqrt(5) + 1) / 2,
  psi_2 = (sqrt(5) - 1) / 2
)
make(plan, trigger = trigger(command = FALSE, depend = FALSE, file = FALSE))
```

Now, even if you wreck all the commands, nothing rebuilds.

```{r}
plan <- drake_plan(
  psi_1 = (sqrt(5) + 1) / 2 + 9999999999999,
  psi_2 = (sqrt(5) - 1) / 2 - 9999999999999
)
make(plan, trigger = trigger(command = FALSE, depend = FALSE, file = FALSE))
```

You can also give different targets to different triggers. Triggers in the `drake` plan override the `trigger` argument to `make()`. Below, `psi_2` always builds, but `psi_1` only builds if it has never been built before.

```{r}
plan <- drake_plan(
  psi_1 = (sqrt(5) + 1) / 2 + 9999999999999,
  psi_2 = target(
    command = (sqrt(5) - 1) / 2 - 9999999999999,
    trigger = trigger(condition = psi_1 > 0)
  )
)
plan
make(plan, trigger = trigger(command = FALSE, depend = FALSE, file = FALSE))
make(plan, trigger = trigger(command = FALSE, depend = FALSE, file = FALSE))
```

Interestingly, `psi_2` now depends on `psi_1`. Since `psi_1` is part of the target `psi_2` because of the **condition** trigger, it needs to be up to date before we attempt `psi_2`. However, since `psi_1` is not part of the command, changing it will not trip the other triggers such as **depend**.

```{r}
vis_drake_graph(plan)
```

In the next toy example below, `drake` reads from a file to decide whether to build `x`. Try it out.

```{r}
plan <- drake_plan(
  x = target(
    1 + 1,
    trigger = trigger(condition = file_in(readRDS("file.rds")))
  )
)
saveRDS(TRUE, "file.rds")
make(plan)
make(plan)
make(plan)
saveRDS(FALSE, "file.rds")
make(plan)
make(plan)
make(plan)
```

In a real project with remote data sources, you may want to use the **condition** trigger to limit your builds to times when enough bandwidth is available for a large download. For example,

```{r, eval = FALSE}
drake_plan(
  x = target(
    command = download_large_dataset(),
    trigger = trigger(condition = is_enough_bandwidth())
  )
)
```

Since the **change** trigger can return any value, it is often easier to use than the **condition** trigger.

```{r}
clean(destroy = TRUE)
plan <- drake_plan(
  x = target(
    command = 1 + 1,
    trigger = trigger(change = sqrt(y))
  )
)
y <- 1
make(plan)
make(plan)
y <- 2
make(plan)
```

In practice, you may want to use the **change** trigger to check a large remote before downloading it.

```{r, eval = FALSE}
drake_plan(
  x = target(
    command = download_large_dataset(),
    trigger = trigger(
      condition = is_enough_bandwidth(),
      change = date_last_modified()
    )
  )
)
```

A word of caution: every non-`NULL` **change** trigger is always evaluated, and its value is carried around in memory throughout `make()`. So if you are not careful, heavy use of the **change** trigger could slow down your workflow and consume extra resources. The **change** trigger should return small values (and should ideally be quick to evaluate). To reduce memory consumption, you may want to return a fingerprint of your trigger value rather than the value itself. See the [`digest`](https://github.com/eddelbuettel/digest) package for more information on computing hashes/fingerprints.

```{r, eval = FALSE}
library(digest)
drake_plan(
  x = target(
    command = download_large_dataset(),
    trigger = trigger(
      change = digest(download_medium_dataset())
    )
  )
)
```

## Alternative trigger modes

Sometimes, you may want to suppress a target without having to worry about turning off every single trigger. That is why the `trigger()` function has a `mode` argument, which controls the role of the **condition** trigger in the decision to build or skip a target. The available trigger modes are `"whitelist"` (default), `"blacklist"`, and `"condition"`.

- `trigger(mode = "whitelist")`: we *rebuild* the target whenever `condition` evaluates to `TRUE`. Otherwise, we defer to the other triggers. This is the default behavior described above in this chapter.
- `trigger(mode = "blacklist")`: we *skip* the target whenever `condition` evaluates to `FALSE`. Otherwise, we defer to the other triggers.
- `trigger(mode = "condition")`: here, the `condition` trigger is the only decider, and we ignore all the other triggers. We *rebuild* target whenever `condition` evaluates to `TRUE` and *skip* it whenever `condition` evaluates to `FALSE`.

## A more practical example

See the ["packages" example](#packages) for a more practical demonstration of triggers and their usefulness.
