# Triggers: decision rules for building targets {#triggers}

```{r setup_triggers, include = FALSE}
suppressMessages(suppressWarnings(library(drake)))
suppressMessages(suppressWarnings(library(tidyverse)))
pkgconfig::set_config("drake::strings_in_dots" = "literals")
clean(destroy = TRUE, verbose = FALSE)
unlink(c("Makefile", "report.Rmd", "shell.sh", "STDIN.o*", "Thumbs.db"))
knitr::opts_chunk$set(
  collapse = TRUE,
  error = TRUE,
  warning = TRUE
)
dat <- system.file(
  file.path("examples", "main", "raw_data.xlsx"),
  package = "drake",
  mustWork = TRUE
)
tmp <- file.copy(from = dat, to = "raw_data.xlsx")
rmd <- system.file(
  file.path("examples", "main", "report.Rmd"),
  package = "drake",
  mustWork = TRUE
)
tmp <- file.copy(from = rmd, to = "report.Rmd")
```

When you call `make()`, `drake` tries to skip as many targets as possible. If it thinks a command will return the same value as last time, it does not bother running it. In other words, `drake` is lazy, and laziness saves you time.

## What are triggers?

To figure out whether it can skip a target, `drake` goes through an intricate checklist of **triggers**:

1. The **missing** trigger: do we already have the target's value from a previous `make()`, or is it missing from the cache?
2. The **command** trigger: did the command in the workflow plan data frame change nontrivially since the last `make()`? Changes to spacing, formatting, and comments are ignored.
3. The **depend** trigger: did any non-file dependencies change since the last `make()`? These could be:
    - Other targets.
    - Imported objects.
    - Imported functions (ignoring changes to spacing, formatting, and comments).
    - Any dependencies of imported functions.
    - Any dependencies of dependencies of imported functions, and so on.
4. The **file** trigger: did any file inputs or file outputs change since the last `make()`? These files are the ones explicitly declared in the command with `file_in()`, `knitr_in()`, and `file_out()`.
5. The **condition** trigger: an optional user-defined piece of code that evaluates to a `TRUE`/`FALSE` value. The target builds if the value is `TRUE`.
6. The **change** trigger:  an optional user-defined piece of code that evaluates to any value (preferably small and quick to compute). The target builds if the value changed since the last `make()`.

If *any* trigger detects something wrong or different with the target or its dependencies, the next `make()` will run the command and (re)build the target.

## Customization

With the `trigger()` function, you can create your own customized checklist of triggers. Let's run a simple workflow with just the **missing** trigger. We deactivate the **command**, **depend**, and **file** triggers by setting the respective `command`, `depend`, and `file` arguments to `FALSE`.

```{r missingtrigger1}
plan <- drake_plan(
  psi_1 = (sqrt(5) + 1) / 2,
  psi_2 = (sqrt(5) - 1) / 2
)
make(plan, trigger = trigger(command = FALSE, depend = FALSE, file = FALSE))
```

Now, even if you wreck all the commands, nothing rebuilds.

```{r missingtrigger2}
plan <- drake_plan(
  psi_1 = (sqrt(5) + 1) / 2 + 9999999999999,
  psi_2 = (sqrt(5) - 1) / 2 - 9999999999999
)
make(plan, trigger = trigger(command = FALSE, depend = FALSE, file = FALSE))
```

You can also give different targets to different triggers. Triggers in the workflow plan data frame override the `trigger` argument to `make()`. Below, `psi_2` always builds, but `psi_1` only builds if the output value is missing beforehand.

```{r missingtrigger3}
plan <- drake_plan(
  psi_1 = (sqrt(5) + 1) / 2 + 9999999999999,
  psi_2 = target(
    command = (sqrt(5) - 1) / 2 - 9999999999999,
    trigger = trigger(condition = psi_1 > 0)
  )
)
plan
make(plan, trigger = trigger(command = FALSE, depend = FALSE, file = FALSE))
make(plan, trigger = trigger(command = FALSE, depend = FALSE, file = FALSE))
```

Interestingly, `psi_2` now depends on `psi_1`. Since `psi_1` is part of the because of the **condition** trigger, it needs to be up to date before we attempt `psi_2`. However, since `psi_1` is not part of the command, changing it will not trip the other triggers such as **depend**.

```{r missingtrigger4}
vis_drake_graph(drake_config(plan))
```

In the next toy example below, `drake` flips a coin to decide whether to build `x`. Try it out.

```{r randomcondition}
set.seed(0)
plan <- drake_plan(
  x = target(
    1 + 1,
    trigger(condition = runif(1) > 0.5)
  )
)
make(plan)
make(plan)
make(plan)
make(plan)
make(plan)
```

In a real project with remote data sources, you may want to use the **condition** trigger to limit your builds to times when enough bandwidth is available for a large download. For example,

```{r tuesday, eval = FALSE}
drake_plan(
  x = target(
    command = download_large_dataset(),
    trigger = trigger(condition = is_enough_bandwidth())
  )
)
```

Since the **change** trigger can return any value, it is often easier to use than the **condition** trigger.

```{r changetrigger}
plan <- drake_plan(
  x = target(
    command = 1 + 1,
    trigger = trigger(change = sqrt(y))
  )
)
y <- 1
make(plan)
make(plan)
y <- 2
make(plan)
```

In practice, you may want to use the **change** trigger to check a large remote before downloading it.

```{r scrapetrigger, eval = FALSE}
drake_plan(
  x = target(
    command = download_large_dataset(),
    trigger = trigger(
      condition = is_enough_bandwidth(),
      change = date_last_modified()
    )
  )
)
```

A word of caution: every non-`NULL` **change** trigger is always evaluated, and its value is always stored in the target's quick-lookup internal metadata (recoverable with `diagnose()`). So if you are not careful, heavy use of the **change** trigger could slow down your workflow. Please take care to make these triggers quick to check, and please keep their return values small. You may want to return a fingerprint of your trigger value rather than the value itself. See the [`digest`](https://github.com/eddelbuettel/digest) package for more information.

```{r scrapetrigger2, eval = FALSE}
library(digest)
drake_plan(
  x = target(
    command = download_large_dataset(),
    trigger = trigger(
      change = digest(download_medium_dataset())
    )
  )
)
```

## A more practical example

See the ["packages" example](#packages) for a more practical demonstration of triggers and their usefulness.

```{r endofline_triggers, echo = FALSE}
clean(destroy = TRUE, verbose = FALSE)
unlink(c("report.Rmd", "raw_data.xlsx", "STDIN.o*", "Thumbs.db"))
```
